---
output:
  html_document:
    keep_md: yes
---

Orbit Catalog for Conjunction Analysis
======================================================

## Synopsis
The first list of TLE objects contains all the ones with ratio-cross section greater that 1.0.
The second list are the ones given by the paper of Rossi et al. on the criticality index (CSI).
A third list of objects are from Max Vasile (small objects).

```{r load_library,results="hide",echo=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(Hmisc)
```

## Raw data

The catalog `tle_latest.txt` is taken form space-track.org:
```{r engine='sh',echo=FALSE,show_tle}
file_name=./tle_latest.txt;
head -6 $file_name ;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $((NUMLIN/3))  (no. of rwos: $NUMLIN )";
```

The catalog `satcat.txt` (containing the physical parameters) comes from celestrak.com:
```{r engine='sh',echo=FALSE,show_satcat}
file_name=./satcat.txt;
head -3 $file_name;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $NUMLIN";
```

An explanation from the variables can be found in the docs folder.

Read file `top15`:
```{r read-top15}
top15 <- tbl_df(read.table(file = "top15", header = TRUE, dec = ".", sep="\t"))
```

Read file `Satellite database.xlsx`:
```{r read-max-datafile}
smallerlist <- tbl_df(read.table("list_small_objects.txt",header = TRUE))
nrow(smallerlist)
```

## Processing data

To process `TLE_LATEST.TXT` into a table containing the Keplerian orbital elements, create a symbolic link or a copy of it and name it `file.fla`.
```{r engine='sh',create_link}
rm -f file.txt
ln -s tle_latest.txt file.txt
```

Then, compile the fortran file `TLE2KEP.f90` (with gfortran, ifort, etc.) and run the executable to transform the 2 lines element catalog in a table format. The output is a file called `file_oe.fla`.
```{r engine='sh',convert_TLE_to_KEP}
gfortran TLE2KEP.f90 -o TLE2KEP
./TLE2KEP
```
```{r engine='sh',echo=FALSE,show_KEP}
file_name=./file_oe.fla;
head -6 $file_name;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $((NUMLIN-1))";
```

To process `SATCAT.TXT` we need change the delimiter of the table (missing values are represented by empty strings and not all can be recognised by R). Compile the fortran code `SAT2TABLE.f90` and run the executable. It will also add an header. The output is a file called `satcat.fla`.
```{r engine='sh',SAT2TABLE}
gfortran SAT2TABLE.f90 -o SAT2TABLE
./SAT2TABLE
```
```{r engine='sh',echo=FALSE,show_new_satcat}
file_name=satcat.fla;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $((NUMLIN-1))";
```

Read the data assigning the right format and merge the two catalogues (here using R)
```{r read_data,cache=FALSE}
tlecat <- tbl_df(read.table(file = "file_oe.fla", header = TRUE, dec = "."))
tlecat
satcat <- tbl_df(read.table(file = "satcat.fla", header = TRUE, dec = ".", sep = "|",na.string=""))
mergecat <- tbl_df(merge(tlecat,satcat,by=c("NORAD_ID")))
```

We are only interseted in a subset of the catalog:
```{r merge_cat}
subcat <- select(mergecat,NORAD_ID,sma_meter,ecc,inc_deg,RAAN_deg,argp_deg,M_deg,RCS)
subcat
```

From these, we select the larger one with RCS>1:
```{r larger-objects}
earth_radius <- 6371
larger <- filter(subcat,RCS>1,sma_meter<(2400+earth_radius)*1000) #,inc_deg<100)
summary(larger)
larger
```

Find the TLE corresponding to the list of objects in the file `list_small_objects.txt`.
```{r find-TLE}
smaller <- tbl_df(merge(subcat,smallerlist,by=c("NORAD_ID")))
smaller <- filter(smaller,RCS<=1)
nrow(smaller)
max(smaller$RCS,na.rm=TRUE)
```


No action needs to be done for the list of objects in the file `top15`.

## Plot

```{r echo=FALSE,fig.height=5,sma_inc}
bigcat <- data.frame(cbind(larger$sma_meter*(1-larger$ecc)/1000 - earth_radius,
                        larger$sma_meter*(1+larger$ecc)/1000 - earth_radius,
                        larger$inc_deg));
names(bigcat) <- c("peri_km","apo_km","inc_deg");

smallcat <- data.frame(cbind(smaller$sma_meter*(1-smaller$ecc)/1000 - earth_radius,
                        smaller$sma_meter*(1+smaller$ecc)/1000 - earth_radius,
                        smaller$inc_deg));
names(smallcat) <- c("peri_km","apo_km","inc_deg");

sc <- data.frame(cbind(top15$a_km*(1-top15$ecc) - earth_radius,
                       top15$a_km*(1+top15$ecc) - earth_radius,
                       top15$inc_deg));
names(sc) <- c("peri_km","apo_km","inc_deg");

with(bigcat, plot(inc_deg, peri_km, col = "blue", pch = 1))
with(bigcat, points(inc_deg, apo_km, col = "red", pch = 1))
with(smallcat, points(inc_deg, peri_km, col = "green", pch = 2))
with(smallcat, points(inc_deg, apo_km, col = "yellow", pch = 2))
with(sc, points(inc_deg, peri_km, col = "black", pch = 20))
```

## Save processed catalog top15 + bigger

We finish writing the catalog `top15` and `larger` into a file with the format need by the Conjunction Algorithm.
```{r writecat1}
orbits1 <- top15[,2:3]
names(orbits1) <- c("a_m","ecc")
orbits1$a_m  <- orbits1$a_m*1000
orbits1$inc  <- top15$inc_deg * pi/360
orbits1$RAAN <- runif(nrow(orbits1),min=0,max=360) * pi/360
orbits1$argp <- runif(nrow(orbits1),min=0,max=360) * pi/360
orbits1$M    <- runif(nrow(orbits1),min=0,max=360) * pi/360

orbits2      <- larger[,2:3]
names(orbits2) <- c("a_m","ecc")
orbits2$inc  <- larger$inc_deg * pi/360
orbits2$RAAN <- larger$RAAN_deg * pi/360
orbits2$argp <- larger$argp_deg * pi/360
orbits2$M    <- larger$M_deg * pi/360

orbit <- rbind(orbits1,orbits2)
orbit <- cbind("",orbit)
write.table(orbit, file = "top15-larger.txt", row.name = FALSE, sep = "\t")
```
The file `top15-larger.txt` has `r nrow(orbit)` rows.

## Save processed catalog larger + smaller
```{r writecat2}
orbits3      <- smaller[,2:3]
names(orbits3) <- c("a_m","ecc")
orbits3$inc  <- smaller$inc_deg * pi/360
orbits3$RAAN <- smaller$RAAN_deg * pi/360
orbits3$argp <- smaller$argp_deg * pi/360
orbits3$M    <- smaller$M_deg * pi/360

orbit <- rbind(orbits2,orbits3)
orbit <- cbind("",orbit)
write.table(orbit, file = "larger-small.txt", row.name = FALSE, sep = "\t")
```
The file `larger-small.txt` has `r nrow(orbit)` rows.