---
output:
  html_document:
    keep_md: yes
---

Satellite Orbit and Decay Analysis
======================================================

## Synopsis
Given the satellite orbits and physical parameters catalogue, we want to select those that are in eccentric orbits and are are going to re-enter the Earth atmosphere. That is the ratio-cross section must be greater that 1.0 and perigee less than 6600 km.

```{r load_library,results="hide"}
library(dplyr)
library(ggplot2)
library(Hmisc)
```

## Raw data

The catalog `tle_latest.txt` is taken form space-track.org:
```{r engine='sh',echo=FALSE,show_tle}
file_name=./tle_latest.txt;
head -6 $file_name ;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $((NUMLIN/3))  (no. of rwos: $NUMLIN )";
```

The catalog `satcat.txt` (containing the physical parameters) comes from celestrak.com:
```{r engine='sh',echo=FALSE,show_satcat}
file_name=./satcat.txt;
head -3 $file_name;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $NUMLIN";
```

An explanation fro the variables can be found in the docs folder.

### Process data
To process `TLE_LATEST.TXT` into a table containing the Keplerian orbital elements, create a symbolic link or a copy of it and name it `file.fla`.
```{r engine='sh',create_link}
rm -f file.txt
ln -s tle_latest.txt file.txt
```

Then, compile the fortran file `TLE2KEP.f90` (with gfortran, ifort, etc.) and run the executable to transform the 2 lines element catalog in a table format. The output is a file called `file_oe.fla`.
```{r engine='sh',convert_TLE_to_KEP}
gfortran TLE2KEP.f90 -o TLE2KEP.x
./TLE2KEP.x
```
```{r engine='sh',echo=FALSE,show_KEP}
file_name=./file_oe.fla;
head -6 $file_name;
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $((NUMLIN-1))";
```


To process `SATCAT.TXT` we need change the delimiter of the table (missing values are represented by empty strings and not all can be recognised by R). Compile the fortran code `SAT2TABLE.f90` and run the executable. It will also add an header. The output is a file called `satcat.fla`.
```{r engine='sh',SAT2TABLE}
gfortran SAT2TABLE.f90 -o SAT2TABLE.x
./SAT2TABLE.x
```
```{r engine='sh',echo=FALSE,show_new_satcat}
file_name=satcat.fla;
head -3 $file_name[,1:8];
NUMLIN=$(cat $file_name | wc -l ); # no. of objects
echo "====================================================================="
echo "Number of objects: $((NUMLIN-1))";
```

Read the data assigning the right format and merge the two catalogues (here using R)
```{r read_data,cache=FALSE}
tlecat <- tbl_df(read.table(file = "file_oe.fla", header = TRUE, dec = "." ))
tlecat
satcat <- tbl_df(read.table(file = "satcat.fla", header = TRUE, dec = ".", sep = "|",na.string=""))
satcat$DECAY <- as.Date(satcat$DECAY)
satcat$LAUNCH <- as.Date(satcat$LAUNCH)
satcat
mergecat <- tbl_df(merge(tlecat,satcat,by=c("NORAD_ID")))
```

We are only interseted in a subset of the catalog:
```{r merge_cat}
subcat <- select(mergecat,NORAD_ID,sma_meter,ecc,inc_deg,RAAN_deg,argp_deg,M_deg,LAUNCH,DECAY,RCS)
subcat
```

Here are some plots.

```{r echo=FALSE,fig.height=5,sma_ecc}
g <- ggplot( subcat, aes( sma_meter/1000, ecc ))
g + geom_point( aes( color = subcat$inc_deg ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "eccentricity")
```

```{r echo=FALSE,fig.height=5,sma_inc}
g <- ggplot( subcat, aes( sma_meter/1000, inc_deg ))
g + geom_point( aes( color = subcat$ecc ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "inclination [deg]")
```

```{r echo=FALSE,fig.height=5,sma_perigee}
g <- ggplot( subcat, aes( log10(sma_meter/1000), log10(sma_meter/1000 * (1 - ecc) )) )
g + geom_point( aes( color = subcat$ecc ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "pericenter [km]") +
    geom_hline( aes(yintercept=seq(log10(7200),log10(6500) ) ) )
```

Let's see whose satellite/debris decay and when, their dimension (cross section) their eccentricity:
```{r filetcat}
sum(!is.na(subcat$DECAY))
subcat_decay <- filter(subcat, !is.na(DECAY))
subcat_decay <- group_by(subcat_decay,DECAY)
select(subcat_decay,DECAY,RCS,ecc)
subcat_decay$quantileRCS <- cut(subcat_decay$RCS,breaks=c(0,0.1,1.0,max(subcat_decay$RCS,na.rm=T)))
levels(subcat_decay$quantileRCS)
subcat_decay$quantileECC <- cut(subcat_decay$ecc,breaks=c(0,0.3,1.0))
levels(subcat_decay$quantileECC)
```

```{r echo=FALSE,fig.height=5,sma_inc_zoom}
g <- ggplot( subcat, aes( sma_meter/1000, inc_deg ))
g + geom_point( aes( color = subcat$ecc ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "inclination [deg]") +
    geom_point( data = subcat_decay, aes( sma_meter/1000, inc_deg ),
               colour="red") +
    coord_cartesian( xlim = c(min(subcat$sma_meter/1000)-1000,50000))
```

```{r echo=FALSE,fig.height=5,sma_perigee_zoom}
g <- ggplot( subcat, aes( log10(sma_meter/1000), sma_meter/1000 * (1 - ecc) ))
g + geom_point( aes( color = subcat$ecc ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "pericenter [km]") +
    #geom_hline( aes(yintercept=seq(7200,6500) ) ) +
    geom_point( data = subcat_decay, aes( log10(sma_meter/1000), sma_meter/1000 * (1 - ecc) ),
               colour="red") +
    coord_cartesian(ylim = c(6400,7300))
```

```{r echo=FALSE,decay_RCS}
g <- ggplot( subcat_decay, aes(x = factor(DECAY)))
g + geom_histogram(aes( fill = factor(quantileRCS))) +
    labs(x = "Date", y = "Number of decays") +
    ggtitle("Decays per date and ratio-cross section") +
    coord_flip()
```

```{r echo=FALSE,decay_ECC}
g <- ggplot( subcat_decay, aes(x = factor(DECAY) ) )
g + geom_histogram(aes( fill = factor(quantileECC))) +
    labs(x = "Date", y = "Number of decays") +
    ggtitle("Decays per date and eccentricity") +
    coord_flip()
```

We are intersted at the big satellite which also decay. So at the end, there are `r sum(subcat_decay$RCS>1)` that satisfy our conditions.
```{r fileterd_cat}
filterd_cat <- filter(subcat_decay,RCS>1)
filterd_cat
```
Of these ones, `r sum( filterd_cat$ecc > 0.3)` are in HEO.

Finally, we look at the orbit with HEO and with perigee less than 6600 km.
```{r filetcat_bis}
dim(subcat)
sum(!is.na(subcat$ecc))
subcat_ecc <- filter(subcat, !is.na(ecc) & ecc > 0.3 & sma_meter/1000*(1-ecc)<6600 )
select(subcat_ecc,sma_meter,ecc,DECAY,RCS)
subcat_ecc
```

```{r echo=FALSE,fig.height=5,sma_inc_zoom_bis}
g <- ggplot( subcat, aes( sma_meter/1000, inc_deg ))
g + geom_point( aes( color = subcat$ecc ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "inclination [deg]") +
    geom_point( data = subcat_ecc, aes( sma_meter/1000, inc_deg ),
               colour="red") +
    coord_cartesian( xlim = c(min(subcat$sma_meter/1000)-1000,50000))
```

```{r echo=FALSE,fig.height=5,sma_perigee_zoom_bis}
g <- ggplot( subcat, aes( sma_meter/1000, sma_meter/1000 * (1 - ecc) ))
g + geom_point( aes( color = subcat$ecc ) )  +
    labs(x = "semimajor axis [km]") +
    labs(y = "pericenter [km]") +
    #geom_hline( aes(yintercept=seq(7200,6500) ) ) +
    geom_point( data = subcat_ecc, aes( sma_meter/1000, sma_meter/1000 * (1 - ecc) ),
               colour="red") +
    coord_cartesian(ylim = c(6400,with(subcat_ecc,max(sma_meter/1000 * (1 - ecc))+100)),
                    xlim = c(min(subcat$sma_meter/1000)-100,max(subcat_ecc$sma_meter/1000)+1000) )
```

```{r echo=FALSE,ecc_RCS}
subcat$quantileRCS <- cut(subcat$RCS,breaks=c(0,0.1,1.0,max(subcat$RCS,na.rm=T)))
levels(subcat$quantileRCS)
subcat$quantileECC <- cut(subcat$ecc,breaks=c(0,0.3,1.0))
levels(subcat$quantileECC)
g <- ggplot( subcat, aes(x = factor(quantileECC) ) )
g + geom_histogram( aes( fill = factor(quantileRCS))) +
    labs(x = "Eccentricity group", y = "Number of objects") +
    ggtitle("Objects divided by eccentricity and ratio-cross section") +
    coord_flip()
```

We are intersted at big satellite in HEO. So at the end, there are `r sum(subcat_ecc$RCS>1)` that satisfy these conditions:

- eccentrycity > 0.3
- perigee < 6600 km
- ratio cross section > 1.0

```{r fileterd_cat_bis}
filterd_cat <- filter(subcat_ecc,RCS>1 )
filterd_cat
write.table(filterd_cat[,1:10], file = "heo_sat.txt", row.name = FALSE, sep = "\t")
```
Of these ones, `r sum( !is.na(filterd_cat$DECAY) )` are expected to decay.

Check the file
```{r check}
tmp <- tbl_df(read.table("heo_sat.txt",sep='\t',header=TRUE))
tmp
```

We finish writing the entire `subcat` data frame into a file:
```{r save_subcat}
write.table(subcat[,1:10], file = "orbitsat.txt", row.name = FALSE, sep = "\t")
```